Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating

### 2017-11-30 SubGenome Analysis.Rmd
Added some captions

```{r pickKSCutoff, echo=FALSE, fig.width=3, fig.height=3}
qplot(
  syntelogs.mutated$ks[syntelogs.mutated$ks < 2.5], 
  geom="histogram", binwidth=.05, na.rm=TRUE
) + geom_vline(
  mapping = NULL, 
  data = NULL, 
  xintercept = 10^params$log10_ks_cutoff, 
  na.rm = FALSE, 
  show.legend = NA
) + labs(
  title="Frequency of KS values by \nsyntenic block (KS < 2.5)", 
  x="KS value of syntenic block", 
  y="Count (# of genes)"
)

@@

qplot(
  log10(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$ks), 
  geom="histogram", 
  binwidth=.01, 
  na.rm=TRUE
) + geom_vline(
  mapping = NULL, 
  data = NULL, 
  xintercept = params$log10_ks_cutoff, 
  na.rm = FALSE, 
  show.legend = NA
) + labs(
  title="Frequency of log10 KS \nvalues", 
  x="log10(KS) value of genes", 
  y="Count (# of genes)"
)
```

@@

```{r reportSyntenicRegions, echo=FALSE, results="asis"}
## Display chromosomes with homeologous genes between org1 and org2
homeologs.chromosome %>%
  select(org_chr1, org_chr2, chromosomeGeneCount) %>%
  spread(org_chr2, chromosomeGeneCount) -> homeologs.spread

@@

# De-emphasize na's by turning them light grey
homeologs.spread.formattedTable <- data.frame(rbind(homeologs.spread[-c(2),], homeologs.spread[c(2),]))
homeologs.spread.formattedTable <- cbind(homeologs.spread.formattedTable[,-c(3, 12)], homeologs.spread.formattedTable[,c(3, 12)])
homeologs.spread.formattedTable <- rename(homeologs.spread.formattedTable, Chromosome=org_chr1)
homeologs.spread.formattedTable[is.na(homeologs.spread.formattedTable)] <- "<font color=\"lightgrey\">na</font>"
homeologs.spread.formattedTable %>% 
  knitr::kable(caption = "Table", 
               align = "lccccccccccr")
```

@@

```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center'}
plot <- graphGenePairDominanceByExperiment(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
plot(plot)
```

@@

```{r subgenome_protein_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center'}
factor <- 4
plot <- graphGenePairDominanceByExperiment(maize.protein.abundance.sample.avg, homeologs.pairs, factor, FALSE) + 
  labs(title = paste0("Which homeolog has greater protein abundance? (factor=",factor,")"))
plot(plot)
```

@@

```{r subgenome_expression_dominance_across_factors, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=12, fig.align='center'}
plot <- graphGenePairDominanceByExperimentAcrossFactors(maize.expression.sample.avg, homeologs.pairs, FALSE)
plot(plot)
```

@@

```{r subgenome_isoform_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=6, fig.height=4, fig.align='center'}
## Plot frequency of isoform counts for each gene
plot <- graphGenePairDominanceByIsoForms(geneTranscript.counts, homeologs.pairs, 4)
plot(plot)
```

@@

```{r subgenome_isoforms_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.align='center'}
# Isoforms
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
names(data)[3] <- "Value_maize1"
names(data)[4] <- "Value_maize2"
# data <- subset(data, data$Value_maize1 > 1 & data$Value_maize2 > 1) ## Ignore single-exon genes
data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)

@@

plot <-
  ggplot(subset(data, Value_maize1 <= 25 & Value_maize2 <= 25), aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(position = "jitter", alpha=.1, aes(color = abs(foldChange))) +
  scale_color_gradient(low = "blue", high = "red") +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  # xlim(0,100) +
  # ylim(0,100) +
  xlim(0,25) +
  ylim(0,25) +
  labs(
    title = paste0("Scatterplot of isoform counts for gene pair(s).\nExcludes ",n," outlier, alpha=.1, foldchange in log2, blue=lm fit."),
    x = "Isoform Counts for Maize1 Gene",
    y = "Isoform Counts for Maize2 Gene"
  ) + 
  # annotate("text",x=85,y=20, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
  annotate("text",x=20,y=3, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
plot(plot)
```

```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, Sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

plot <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = Sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  scale_x_log10() + scale_y_log10() +
  labs(
    title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1, reference line"),
    x = "Expression for Maize1 Gene (log scale)",
    y = "Expression for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~Sample, nrow=6)
plot(plot)
```

```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID")) %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, Sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

plot <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = Sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  scale_x_log10() + scale_y_log10() +
  labs(
    title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1, reference line"),
    x = "Abundance for Maize1 Gene (log scale)",
    y = "Abundance for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~Sample, nrow=6)
plot(plot)
```

>>>>

```{r pickKSCutoff, echo=FALSE, fig.width=3, fig.height=3, fig.cap = cap}
qplot(
  syntelogs.mutated$ks[syntelogs.mutated$ks < 2.5], 
  geom="histogram", binwidth=.05, na.rm=TRUE
) + geom_vline(
  mapping = NULL, 
  data = NULL, 
  xintercept = 10^params$log10_ks_cutoff, 
  na.rm = FALSE, 
  show.legend = NA
) + labs(
  title="Frequency of KS values by \nsyntenic block (KS < 2.5)", 
  x="KS value of syntenic block", 
  y="Count (# of genes)"
)

@@

qplot(
  log10(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$ks), 
  geom="histogram", 
  binwidth=.01, 
  na.rm=TRUE
) + geom_vline(
  mapping = NULL, 
  data = NULL, 
  xintercept = params$log10_ks_cutoff, 
  na.rm = FALSE, 
  show.legend = NA
) + labs(
  title="Frequency of log10 KS \nvalues", 
  x="log10(KS) value of genes", 
  y="Count (# of genes)"
)
cap <- "Fig 2.1: "
```

@@

```{r reportSyntenicRegions, echo=FALSE, results="asis", fig.cap = cap}
## Display chromosomes with homeologous genes between org1 and org2
homeologs.chromosome %>%
  select(org_chr1, org_chr2, chromosomeGeneCount) %>%
  spread(org_chr2, chromosomeGeneCount) -> homeologs.spread

@@

# De-emphasize na's by turning them light grey
homeologs.spread.formattedTable <- data.frame(rbind(homeologs.spread[-c(2),], homeologs.spread[c(2),]))
homeologs.spread.formattedTable <- cbind(homeologs.spread.formattedTable[,-c(3, 12)], homeologs.spread.formattedTable[,c(3, 12)])
homeologs.spread.formattedTable <- rename(homeologs.spread.formattedTable, Chromosome=org_chr1)
homeologs.spread.formattedTable[is.na(homeologs.spread.formattedTable)] <- "<font color=\"lightgrey\">na</font>"
homeologs.spread.formattedTable %>% 
  knitr::kable(caption = "Table", 
               align = "lccccccccccr")
cap <- "Table 3.1: "
```

@@

```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
plot <- graphGenePairDominanceByExperiment(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
plot(plot)
cap <- "Fig 9.1:"
```

@@

```{r subgenome_protein_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
factor <- 4
plot <- graphGenePairDominanceByExperiment(maize.protein.abundance.sample.avg, homeologs.pairs, factor, FALSE) + 
  labs(title = paste0("Which homeolog has greater protein abundance? (factor=",factor,")"))
plot(plot)
cap <- "Fig 9.2:"
```

@@

```{r subgenome_expression_dominance_across_factors, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=12, fig.align='center', fig.cap = cap}
plot <- graphGenePairDominanceByExperimentAcrossFactors(maize.expression.sample.avg, homeologs.pairs, FALSE)
plot(plot)
cap <- "Fig 9.3:"
```

@@

```{r subgenome_isoform_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=6, fig.height=4, fig.align='center', fig.cap = cap}
## Plot frequency of isoform counts for each gene
plot <- graphGenePairDominanceByIsoForms(geneTranscript.counts, homeologs.pairs, 4)
plot(plot)
cap <- "Fig 9.4: "
```

@@

```{r subgenome_isoforms_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.align='center', fig.cap = cap}
# Isoforms
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
names(data)[3] <- "Value_maize1"
names(data)[4] <- "Value_maize2"
# data <- subset(data, data$Value_maize1 > 1 & data$Value_maize2 > 1) ## Ignore single-exon genes
data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)

@@

plot <-
  ggplot(subset(data, Value_maize1 <= 25 & Value_maize2 <= 25), aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(position = "jitter", alpha=.1, aes(color = abs(foldChange))) +
  scale_color_gradient(low = "blue", high = "red") +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  # xlim(0,100) +
  # ylim(0,100) +
  xlim(0,25) +
  ylim(0,25) +
  labs(
    title = paste0("Scatterplot of isoform counts for gene pair(s).\nExcludes ",n," outlier, alpha=.1, foldchange in log2, blue=lm fit."),
    x = "Isoform Counts for Maize1 Gene",
    y = "Isoform Counts for Maize2 Gene"
  ) + 
  # annotate("text",x=85,y=20, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
  annotate("text",x=20,y=3, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
plot(plot)

cap <- "Fig 9.5: "
```

```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, Sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

plot <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = Sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  scale_x_log10() + scale_y_log10() +
  labs(
    title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1, reference line"),
    x = "Expression for Maize1 Gene (log scale)",
    y = "Expression for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~Sample, nrow=6)
plot(plot)

cap <- "Fig 9.5: "
```

```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID")) %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, Sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

plot <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = Sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  scale_x_log10() + scale_y_log10() +
  labs(
    title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1, reference line"),
    x = "Abundance for Maize1 Gene (log scale)",
    y = "Abundance for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~Sample, nrow=6)
plot(plot)

cap <- "Fig 9.6: "
```
